Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-13 WSSSPE3/report/WSSSPE3_review.tex
changes to changes, some undoing

\item \textbf{Jeffrey Carver: \textit{On the Need for Software Engineering
Support for Sustainable Scientific Software}} Carver argued that for scientific
software to be truly sustainable, there is a need for developers to use
appropriate software engineering practices. His experience interacting with
scientific teams indicates that choosing and tailoring these practices is not a
trivial exercise. There is a general culture clash between software engineering
and science that hinders our ability to communicate and choose appropriate
methods. In addition, many experienced scientific software developers appear to
be unaware of software engineering practices that may be beneficial to them. The
most appropriate software engineering practices are those that are lightweight,
properly tailored, and focus on the key software development problems faced by
scientists. In order to increase the use of software engineering in science, we
need more documented success stories. These successes need to be socialized 
\choinote{not sure if ``socialize'' can be used in such a way}
within the scientific community through workshops like the Software Engineering
for Science Workshop
Series\footnote{\href{http://www.SE4Science.org/workshops}{http://www.SE4Science.org/workshops}}
and the new Software Engineering track in the \textit{Computing in Science and
Engineering} magazine.

@@

\item \textbf{Stan Ahalt, Bruce Berriman, Maxine Brown, Jeffrey Carver, Neil
Chue Hong, Allison Fish, Ray Idaszak, Greg Newman, Dhabaleswar Panda, Abani
Patra, Elbridge Gerry Puckett, Chris Roland, Douglas Thain, Selcuk Uluagac, and
Bo Zhang: \textit{Scientific Software Success: Developing Metrics While
Developing Community}} The effort behind this talk by Ray Idaszak started from a
break-out group at an NSF SI2 workshop in 2015, and centerd around creating a
framework for creation of metrics for scientific software. This framework would
improve both the metrics and the software it evaluates, and could also serve as a tool for building
a community around the idea. With especially the latter idea (building
a community) in mind, a software ``peer review group'' would be created, representing stakeholders who
will self-review software created by their respective communities, and will
concurrently develop metrics. The whole project should be community-governed,
without a single institution overseeing the activities or infrastructure, with the
hope to evolving community-generated and adopted standards. The generation of
metrics would be tied to the actual evaluation of software, creating an
incentive by improving the evaluated software itself during this process. The
framework code would provide infrastructure for the creation of metrics and
evaluation, and forums for generation of software success metrics. It would also
support code reviews of the evaluated software. An open question is whether it
is possible to fit the resulting metrics in a common template. So far, this is
still in a design phase, with a white paper at the CSESSP workshop 2015 and this
talk. The WSSSPE workshop is seen as a forum for the community to assemble and
act, and is planned to be used also in the future to build this community and
framework.

>>>>

\item \textbf{Jeffrey Carver: \textit{On the Need for Software Engineering
Support for Sustainable Scientific Software}} Carver argued that for scientific
software to be truly sustainable, there is a need for developers to use
appropriate software engineering practices. His experience interacting with
scientific teams indicates that choosing and tailoring these practices is not a
trivial exercise. There is a general culture clash between software engineering
and science that hinders our ability to communicate and choose appropriate
methods. In addition, many experienced scientific software developers appear to
be unaware of software engineering practices that may be beneficial to them. The
most appropriate software engineering practices are those that are lightweight,
properly tailored, and focus on the key software development problems faced by
scientists. In order to increase the use of software engineering in science, we
need more documented success stories. These successes need to be socialized 
\choinote{not sure if ``socialize'' can be used in such a way}
within the scientific community through workshops like the Software Engineering
for Science workshop
series\footnote{\href{http://www.SE4Science.org/workshops}{http://www.SE4Science.org/workshops}}
and the new Software Engineering track in \textit{Computing in Science and
Engineering} magazine.

@@

\item \textbf{Stan Ahalt, Bruce Berriman, Maxine Brown, Jeffrey Carver, Neil
Chue Hong, Allison Fish, Ray Idaszak, Greg Newman, Dhabaleswar Panda, Abani
Patra, Elbridge Gerry Puckett, Chris Roland, Douglas Thain, Selcuk Uluagac, and
Bo Zhang: \textit{Scientific Software Success: Developing Metrics While
Developing Community}} The effort behind this talk by Ray Idaszak started from a
break-out group at an NSF SI2 workshop in 2015, and centers around creating a
framework for creation of metrics for scientific software. This framework would
improve both the metrics and the software it evaluates, and could also serve as a tool for building
a community around the idea. With especially the latter idea (building
a community) in mind, a software ``peer review group'' would be created, representing stakeholders who
will self-review software created by their respective communities, and will
concurrently develop metrics. The whole project should be community-governed,
without a single institution overseeing the activities or infrastructure, with the
hope to evolving community-generated and adopted standards. The generation of
metrics would be tied to the actual evaluation of software, creating an
incentive by improving the evaluated software itself during this process. The
framework code would provide infrastructure for the creation of metrics and
evaluation, and forums for generation of software success metrics. It would also
support code reviews of the evaluated software. An open question is whether it
is possible to fit the resulting metrics in a common template. So far, this is
still in a design phase, with a white paper at the CSESSP workshop 2015 and this
talk, but the WSSSPE workshops are seen as a forum for the community to assemble and
act, and is planned to be used also in the future to build this community and
framework.
